---
title: "The Historical Microsoft Analysis"
author: "Milica Cudina"
date: "`r Sys.Date()`"
output: pdf_document
---
<!-- The author of this template is Dr. Gordan Zitkovic.-->
<!-- The code chunk below contains some settings that will  -->
<!-- make your R code look better in the output pdf file.  -->
<!-- If you are curious about what each option below does, -->
<!-- go to https://yihui.org/knitr/options/ -->
<!-- If not, feel free to disregard everything ...  -->
```{r echo=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.align="center",
  fig.pos="t",
  strip.white = TRUE
)
```
<!-- ... down to here. -->

---

## Task 1. 
First, we will read the data from our csv file "MSFT.csv" into a data.frame called `msft`:

```{r}
msft<-read.csv("MSFT.csv")
```

## Task 2.
Draw the time-plot of the evolution of the closing stock price (not the adjusted). You do **not** need to put the calendar days on the horizontal axis, but you **do** need to label your axes and give your time-plot a title indicating the dates.   

```{r}
close.price=msft$Close
#close.price
plot(close.price, 
     main="Daily closing prices of Microsoft over the last year",
     xlab="Time",
     ylab="Close price",
     pch=20, 
     type="l",
     col="blue")
```

## Task 3. 
What are the mean and the standard deviation of the prices?

```{r}
mean(close.price)
sd(close.price)
#just for laughs
#summary(close.price)
#hist(close.price, breaks=25)
```

## Task 4. 
The **simple daily return** of the stock over a day indexed by $t$ is defined as 
$$
\frac{\text{price at end of day $t$}-\text{price at end of day $(t-1)$}}{\text{price at end of day $(t-1)$}} 
$$
Construct the vector of simple daily returns over the last year. Provide a visualization of the returns. What can you say about the characteristics of the distribution based on the above plot?

```{r}
close.price.end=close.price[-1]
close.price.beg=close.price[-length(close.price)]
returns=(close.price.end-close.price.beg)/close.price.beg
#returns
hist(returns, breaks=25, 
     main="Simple Daily Returns of MSFT",
     xlab="Returns",
     col="bisque")
#outliers?
#boxplot(returns)
```
**Kind of symmetric; kind of unimodal.**

## Task 5. 
What are the mean and the standard deviation of the returns?

```{r}
mu=mean(returns)
mu
sigma=sd(returns)
sigma
```

## Task 6. 
How would you **fit** a distribution to the above histogram? Let's 

* propose a model for the returns; 
* calculate the point estimates of the parameters of the model;
* superimpose the density of the modelling distribution onto the histogram. 

```{r}
hist(returns, breaks=25, 
     main="Simple Daily Returns of MSFT",
     xlab="Returns",
     col="bisque",
     prob=TRUE)
curve(dnorm(x, mean=mu, sd=sigma), col="red", lwd=2, add=TRUE)
```


## Task x. 
What else could we ask? 
