---
title: "The Historical Microsoft Analysis"
author: "Milica Cudina"
date: "`r Sys.Date()`"
output: pdf_document
---
<!-- The author of this template is Dr. Gordan Zitkovic.-->
<!-- The code chunk below contains some settings that will  -->
<!-- make your R code look better in the output pdf file.  -->
<!-- If you are curious about what each option below does, -->
<!-- go to https://yihui.org/knitr/options/ -->
<!-- If not, feel free to disregard everything ...  -->
```{r echo=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.align="center",
  fig.pos="t",
  strip.white = TRUE
)
```
<!-- ... down to here. -->

---

## Task 1. 
First, we will read the data from our csv file "MSFT.csv" into a data.frame called `msft`:

```{r}
msft<-read.csv("MSFT.csv")
msft
#View(msft)
```


## Task 2.
Draw the time-plot of the evolution of the closing stock price (not the adjusted). You do **not** need to put the calendar days on the horizontal axis, but you **do** need to label your axes and give your time-plot a title indicating the dates.   

```{r}
close.price<-msft$Close
l=length(close.price)

plot(close.price, 
     type="l", col="blue", lwd=2, 
     xlab="Time", ylab="Close price", 
     main="Historical MSFT Prices")
```

## Task 3. 
What are the mean and the standard deviation of the prices?

```{r}
mean(close.price)
sd(close.price)
#just for fun
summary(close.price)
```


## Task 4. 
The **simple daily return** of the stock over a day indexed by $t$ is defined as 
$$
\frac{\text{price at end of day $t$}-\text{price at end of day $(t-1)$}}{\text{price at end of day $(t-1)$}} 
$$
Construct the vector of simple daily returns over the last year. Provide a visualization of the returns. What can you say about the characteristics of the distribution based on the above plot?

```{r}
ends=close.price[-1]
begs=close.price[-l]
returns=(ends-begs)/begs
#returns

#plotting a histogram
hist(returns, breaks=20)
```


## Task 5. 
What are the mean and the standard deviation of the returns?

```{r}
#mean of daily returns
(mu.hat=mean(returns))
#standard deviation of daily returns
(sigma.hat=sd(returns))

#just for fun
#annual mean return?
(252*mu.hat)
#annual volatility 
(sqrt(252)*sigma.hat)
```


## Task 6. 
How would you **fit** a distribution to the above histogram? Let's 

* propose a model for the returns; 
* calculate the point estimates of the parameters of the model;
* superimpose the density of the modelling distribution onto the histogram. 

```{r}
#plotting a histogram
hist(returns, breaks=20, 
     main="Model for MSFT returns", 
     col="lavender", xlab="Returns", prob=TRUE)

curve(dnorm(x, mean=mu.hat, sd=sigma.hat), add=TRUE, 
      lwd=2, col="purple")

```


## Task x. 
What else could we ask? 
