---
title: "The Duke GPA Data Analysis: In Class"
author: "the class"
date: "`r Sys.Date()`"
output: pdf_document
---
<!-- The author of this template is Dr. Gordan Zitkovic.-->
<!-- The code chunk below contains some settings that will  -->
<!-- make your R code look better in the output pdf file.  -->
<!-- If you are curious about what each option below does, -->
<!-- go to https://yihui.org/knitr/options/ -->
<!-- If not, feel free to disregard everything ...  -->
```{r echo=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.align="center",
  fig.pos="t",
  strip.white = TRUE
)
```
<!-- ... down to here. -->

---

## Task 1. 
First, we will read the data from our csv file "gpa.csv" into a data.frame called `gpa.data`:
```{r}
gpa.data<-read.csv("gpa.csv")
#gpa.data
```

If you want to see what your data.frame looks like, you can click on it in the **Global environment** in the upper right pane. The data.frame will get displayed in the upper left pane. 

## Task 2.
You interested in the types and names of the variables in your data.frame. What do you run?
```{r}
ls.str(gpa.data)
```

You see that the students/**cases** all have corresponding **rows**. They are labeled by the row indices. The **column** names stand for the variable names.

Then, you can do a bit of exploratory analysis. 

## Task 3. 
What are the minimum and maximum GPAs? What is the mean GPA? What is the median GPA?

```{r}
gpa<-gpa.data$gpa
min(gpa)
max(gpa)
mean(gpa)
sd(gpa)
summary(gpa)
```


## Task 4. 
Plot the histogram of the GPAs. Make sure that your plot has the main title and that the axes are also labeled. 

```{r}
hist(gpa,
     breaks=15,
     main="Histogram of GPA",
     xlab="GPA",
     ylab="Frequency",
     col="lightblue")
```

## Task 5. 
Plot the histogram of the number of hours spent studying per week. Make sure that your plot has the main title and that the axes are also labeled. 

```{r}
hrs<-gpa.data$studyweek
hist(hrs,
     breaks=18,
     main="Histogram of hours spent studying",
     xlab="Hours",
     ylab="Frequency",
     col="steelblue")
```

## Task 6. 
Is the mean number of hours spent studying different for females than for males? 

```{r}
gpa.data$gender
gpa.data$gender=="female"
hrs.f=hrs[gpa.data$gender=="female"]
hrs.f
hrs.m=hrs[gpa.data$gender=="male"]
hrs.m
mean(hrs.f)-mean(hrs.m)
```


## Task 7. 
Any difference in the GPA?
```{r}
gpa.f=gpa[gpa.data$gender=="female"]
gpa.m=gpa[gpa.data$gender=="male"]
mean(gpa.f)-mean(gpa.m)
```

## Task 8. 
Is there a relationship between the hours studied and the GPA?

```{r}
plot(hrs,gpa, pch=19, col="blue",
     main="Scatterplot of hours studied and GPA",
     xlab="Hours",
     ylab="GPA")
slr=lm(gpa ~ hrs)
slr
summary(slr)
abline(slr, col="red")
```

## Task x. 
What else could we ask? 
```{r}
plot(gpa.data$studyweek,gpa.data$gpa, pch=16, 
     col=as.factor(gpa.data$gender))
legend(40, 4.75, legend=c("Male", "Female"),
       col=c("red", "black"), pch=16, cex=0.8)
```
