---
title: "Problem Set #1: In Class"
author: "Milica Cudina"
date: "`r Sys.Date()`"
output: pdf_document
---
<!-- The author of this template is Dr. Gordan Zitkovic.-->
<!-- The code chunk below contains some settings that will  -->
<!-- make your R code look better in the output pdf file.  -->
<!-- If you are curious about what each option below does, -->
<!-- go to https://yihui.org/knitr/options/ -->
<!-- If not, feel free to disregard everything ...  -->
```{r echo=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.align="center",
  fig.pos="t",
  strip.white = TRUE
)
```
<!-- ... down to here. -->

---

Create an `R`-notebook which prints out the answers to the following problems. All of your work and calculations **must** be done in `R`.

# Arithmetic (1 point each)
## Problem 1. 
Divide 792 by 11. 

*Solution:*
```{r}
792/11
```

## Problem 2. 
Find the square root of $\pi$

```{r}
sqrt(pi)
```


## Problem 3. 
Calculate $e^2$

```{r}
exp(2)
```


## Problem 4. 
Find the fourth root of 256.

```{r}
256^(1/4)
sqrt(sqrt(256))
```


## Problem 5. 
Calculate the natural logarithm of 2.

```{r}
log(2)
#just for laughs
log2(2)
```


# Variables and vectors
## Problem 6. **(2 points)** 
Assign the value 3 to variable `a` and  the value 7 to variable `b`; print out their product

```{r}
a=3
b<-7
a*b
```

## Problem 7. **(3 points)** 
Set the vector `v` to be the numbers between -2 and 4 (inclusive); set the vector `w` to be the first 7 natural numbers; what do you get when you multiply `v` by `w`?

```{r}
v=seq(-2, 4, 1)
v
w=1:7
w
v*w
(v*w)
print(v*w)
```


# Loops
## Problem 8. **(5 points)**
Write and execute a chunk of code which calculates and prints out the sum of squares of the first $n$ natural numbers for $n=1, 2, \dots, 25$. 

*Solution:*
This is the sum of the first $10$ natural numbers: 

```{r}
x=1:10
x
y=x^2
y
sum(y)
sum((1:10)^2)
```

The full solution would be:

```{r}
for (n in 1:25){
  print(sum((1:n)^2))
}
```


