---
title: "The Historical Microsoft Analysis"
author: "Milica Cudina"
date: "`r Sys.Date()`"
output: pdf_document
---
<!-- The author of this template is Dr. Gordan Zitkovic.-->
<!-- The code chunk below contains some settings that will  -->
<!-- make your R code look better in the output pdf file.  -->
<!-- If you are curious about what each option below does, -->
<!-- go to https://yihui.org/knitr/options/ -->
<!-- If not, feel free to disregard everything ...  -->
```{r echo=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.align="center",
  fig.pos="t",
  strip.white = TRUE
)
```
<!-- ... down to here. -->

---

## Task 1. 
First, we will read the data from our csv file "MSFT.csv" into a data.frame called `msft`:

```{r}
msft<-read.csv("MSFT.csv")
msft
ls.str(msft)
```


## Task 2.
Draw the time-plot of the evolution of the closing stock price (not the adjusted). You do **not** need to put the calendar days on the horizontal axis, but you **do** need to label your axes and give your time-plot a title indicating the dates.   

```{r}
close=msft$Close
length(close)
plot(close, type="l", col="blue",
     main="Daily closing prices of Msft",
     xlab="Time",
     ylab="Stock price")
```

## Task 3. 
What are the mean and the standard deviation of the prices?

```{r}
mean(close)
sd(close)
```

## Task 4. 
The **simple daily return** of the stock over a day indexed by $t$ is defined as 
$$
\frac{\text{price at end of day $t$}-\text{price at end of day $(t-1)$}}{\text{price at end of day $(t-1)$}} 
$$
Construct the vector of simple daily returns over the last year. Provide a visualization of the returns. What can you say about the characteristics of the distribution based on the above plot?

```{r}
close.end=close[-1]
close.beg=close[-length(close)]
returns=(close.end-close.beg)/close.beg
#plot(returns, type="l")
hist(returns, breaks=20, 
     main="Histogram of daily returns",
     xlab="Returns",
     col="bisque")
```

## Task 5. 
What are the mean and the standard deviation of the returns?

```{r}
mu=mean(returns)
mu
sd=sd(returns)
sd
```

## Task 6. 
How would you **fit** a distribution to the above histogram? Let's 

* propose a model for the returns; 
* calculate the point estimates of the parameters of the model;
* superimpose the density of the modelling distribution onto the histogram. 

```{r}
hist(returns, breaks=20, 
     main="Histogram of daily returns",
     xlab="Returns",
     col="bisque", 
     prob=TRUE)
curve(dnorm(x, mean=mu, sd=sd), lwd=2, col="red", add=TRUE)
```

## Task x. 
What else could we ask? 

```{r}
plot(close[-length(close)], returns, pch=20, col="blue")
lm=lm(returns ~ close[-length(close)])
abline(lm)
```

