---
title: "Quiz #7"
author: "Inigo Montoya"
date: "`r Sys.Date()`"
output: pdf_document
---
<!-- The author of this template is Dr. Gordan Zitkovic.-->
<!-- The code chunk below contains some settings that will  -->
<!-- make your R code look better in the output pdf file.  -->
<!-- If you are curious about what each option below does, -->
<!-- go to https://yihui.org/knitr/options/ -->
<!-- If not, feel free to disregard everything ...  -->
```{r echo=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.align="center",
  fig.pos="t",
  strip.white = TRUE
)
```
<!-- ... down to here. -->

---

Create an `R`-notebook which prints out the answers to the following problems. Knit the Rmd file into a pdf. Upload the pdf of your solutions onto Canvas. All of your work and calculations **must** be done in `R`.

---

# Arithmetic (1 point each)
## Problem 1. 
Divide 385 by 11. 
```{r}
385/11
```


## Problem 2. 
Find the square root of $2\pi$.

## Problem 3. 
Calculate the square root of $e$.

## Problem 4. 
Find the third root of 27.

## Problem 5. 
Calculate the natural logarithm of 10.

# Variables and vectors
## Problem 6. **(2 points)** 
Assign the value 7 to variable `a` and  the value 4 to variable `b`; print out their product

## Problem 7. **(3 points)** 
Set the vector `v` to be the numbers between -1 and 4 (inclusive); set the vector `w` to be the first 6 natural numbers, i.e., integers strictly greater than zero; what do you get when you multiply `v` by `w`?

# Loops
## Problem 8. **(5 points)**
Write and execute a chunk of code which calculates and prints out the sum of squares of the first $n$ natural numbers for $n=1, 2, \dots, 25$. 

**Be extra careful: The number of terms in the sum is varying from $1$ to $25$; you need to produce code which outputs $25$ numbers.**
