---
title: "Problem set #8: Binomial Monte Carlo: Solutions"
author: "Milica Cudina"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

Let the **volatility** of a stock be the standard deviation of its (continuously compounded) realized return on an annual basis. Then, we can define the up and down factors in the so-called *forward binomial tree* for a **non-dividend-paying** stock as
\begin{equation}
\begin{split}
u & = e^{rh + \sigma\sqrt{h}}\\
d & = e^{rh - \sigma\sqrt{h}}\\
\end{split}
\end{equation}

Let the continuously compounded, risk-free interest rate be $0.04$. 

```{r}
r=0.04
```

Consider a stock whose current price is \$100 and whose volatility is $0.25$. We will be pricing a one-year, at-the-money call option in a variety of ways here. 
```{r}
s0=100
sigma=0.25
T=1
K=100
```


## Problem #1: Analytic one period
Price the option above using a one period binomial tree. 

*Solution:*
```{r}
n=1
h=T/n
u.1=exp(r*h+sigma*sqrt(h))
d.1=exp(r*h-sigma*sqrt(h))
s.u=s0*u.1
s.d=s0*d.1

#the payoff function
v.c<-function(x){
  max(x-K, 0)
}

p.star=(exp(r*h)-d.1)/(u.1-d.1)
v.u=v.c(s.u)
v.d=v.c(s.d)

v.0=exp(-r*T)*(p.star*v.u+(1-p.star)*v.d)
v.0
```

## Problem #2: Monte Carlo one period
Price the option above using Monte Carlo a one period binomial tree. Use $10000$ simulations. 

*Solution:*
```{r}
n.sims=10000
factors=c(u.1, d.1)
final.prices=s0*sample(factors, size=n.sims, prob=c(p.star, 1-p.star), replace=TRUE)
final.prices
payoffs<-pmax(final.prices-K,0)
payoffs
v.bar=mean(payoffs)
v.0.mc<-exp(-r*T)*v.bar
v.0.mc
```

## Problem #3: Analytic two periods
Price the above option using a two-period binomial tree. 

*Solution:*
```{r}
n=2
h=T/n
u=exp(r*h+sigma*sqrt(h))
d=exp(r*h-sigma*sqrt(h))
s.T=s0*c(u^2, u*d, d^2)
s.T

p.star=(exp(r*h)-d)/(u-d)
v.T=pmax(s.T-K,0)
v.T

v.0=exp(-r*T)*(v.T[1]*dbinom(2, size=2, p.star)+
                 v.T[2]*dbinom(1, size=2, p.star)+
                 v.T[3]*dbinom(0, size=2, p.star))
v.0
```

## Problem #4: Monte Carlo two periods
Price the option above using Monte Carlo a two period binomial tree. Use $10000$ simulations. 

*Solution:*
```{r}
x=rbinom(n.sims, size=2, prob=p.star)
s.T=s0*u^x*d^(2-x)
s.T

v.T=pmax(s.T-K,0)
v.T

v.bar=mean(v.T)
v.0.mc=exp(-r*T)*v.bar
v.0.mc
```

## Problem #5: Analytic one hundred periods
Price the above option using a $100-$period binomial tree. 

*Solution:*
```{r}
n=100
h=T/n
u=exp(r*h+sigma*sqrt(h))
d=exp(r*h-sigma*sqrt(h))
k=n:0
s.T=s0*u^k*d^(n-k)
s.T
barplot(s.T)

p.star=(exp(r*h)-d)/(u-d)
v.T=pmax(s.T-K,0)
v.T

rn.probs=dbinom(k, size=n, p.star)
rn.probs

v.0=exp(-r*T)*sum(v.T*rn.probs)
v.0
```

## Problem #6: Monte Carlo with one hundred periods
Price the option above using Monte Carlo with a hundred period binomial tree. Use $10000$ simulations. 

*Solution:*
```{r}
periods=100
x=rbinom(n.sims, size=periods, prob=p.star)
s.T=s0*u^x*d^(periods-x)
s.T

v.T=pmax(s.T-K,0)
v.T

v.bar=mean(v.T)
v.0.mc=exp(-r*T)*v.bar
v.0.mc
```
