---
title: "Problem set #11: Binomial Monte Carlo"
author: "Milica Cudina"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

Let the **volatility** of a stock be the standard deviation of its (continuously compounded) realized return on an annual basis. Then, we can define the up and down factors in the so-called *forward binomial tree* for a **non-dividend-paying** stock as
\begin{equation}
\begin{split}
u & = e^{rh + \sigma\sqrt{h}}\\
d & = e^{rh - \sigma\sqrt{h}}\\
\end{split}
\end{equation}

Let the continuously compounded, risk-free interest rate be $0.04$. 

```{r}
r=0.04
```

Consider a stock whose current price is \$100 and whose volatility is $0.25$. We will be pricing a one-year, at-the-money call option in a variety of ways here. 

```{r}
#about the stock 
s0=100
sigma=0.25

#about the call
T=1
K=s0
```


## Problem #1: Analytic one period
Price the option above using a one period binomial tree. 

```{r}
# number of periods
n=1

# length of one period
h=T/n

# define the factors in our forward binomial tree
(u=exp(r*h+sigma*sqrt(h)))
(d=exp(r*h-sigma*sqrt(h)))

# the risk-neutral probability
(p=(exp(r*h)-d)/(u-d))

# the forward-tree shortcut
(p.tilde=1/(1+exp(sigma*sqrt(h))))

# the possible stock prices
(s.u=u*s0)
(s.d=d*s0)

# the possible payoffs
(v.u=max(s.u-K,0))
(v.d=max(s.d-K,0))

# the initial option price
(v.0=exp(-r*T)*(v.u*p+v.d*(1-p)))
```


## Problem #2: Monte Carlo one period
Price the option above using Monte Carlo a one period binomial tree. Use $10000$ simulations. 

```{r}
# the number of simulations
n.sim=10^5

# possible results of a cointoss
factors=c(u,d)
# their risk-neutral probabilities
probs=c(p, 1-p)

# simulated stock prices
s.T=s0*sample(factors, size=n.sim, prob=probs, replace=TRUE)

# simulated payoffs
v.T=pmax(s.T-K,0) # caveat: it MUST be the pointwise max
v.T

# average payoff
v.bar=mean(v.T)

# discounting 
(v.0.mc=exp(-r*T)*v.bar)
```


## Problem #3: Analytic two periods
Price the above option using a two-period binomial tree. 

```{r}
# number of periods
n=2

# length of one period
h=T/n

# define the factors in our forward binomial tree
(u=exp(r*h+sigma*sqrt(h)))
(d=exp(r*h-sigma*sqrt(h)))

# the risk-neutral probability
(p=(exp(r*h)-d)/(u-d))

# possible stock prices
(s.T=c(s0*u^2, s0*u*d, s0*d^2))

# possible payoffs
(v.T=pmax(s.T-K,0))

# the risk-neutral probabilities of reaching the final nodes
(probs=dbinom(c(2,1,0), size=n, prob=p))

# expected risk-neutral payoff
(sum(probs*v.T))

# discounting
(v.0=exp(-r*T)*sum(probs*v.T))
```


## Problem #4: Monte Carlo two periods
Price the option above using Monte Carlo a two period binomial tree. Use $10000$ simulations. 

```{r}
# steppingstone
# simulating the heights of the final stock prices
x=rbinom(n.sim, n, p)

# simulated stock prices
s.T=s0*u^x*d^(n-x)

# simulated payoffs
v.T=pmax(s.T-K,0) # caveat: it MUST be the pointwise max
#v.T

# average payoff
v.bar=mean(v.T)

# discounting 
(v.0.mc=exp(-r*T)*v.bar)
```


## Problem #5: Analytic one hundred periods
Price the above option using a $100-$period binomial tree. 

```{r}
# number of periods
n=100

# length of one period
h=T/n

# define the factors in our forward binomial tree
(u=exp(r*h+sigma*sqrt(h)))
(d=exp(r*h-sigma*sqrt(h)))

# the risk-neutral probability
(p=(exp(r*h)-d)/(u-d))

# the possible stock prices
k=n:0
s.T=s0*u^k*d^(n-k)

# the possible payoffs
v.T=pmax(s.T-K,0)

# their respective probabilities
probs=dbinom(k, size=n, prob=p)

# the risk-neutral expectation
sum(probs*v.T)

# discounting
(v.0=exp(-r*T)*sum(probs*v.T))
```

## Problem #6: Monte Carlo with one hundred periods
Price the option above using Monte Carlo with a hundred period binomial tree. Use $10000$ simulations. 

```{r}
# steppingstone
# simulating the heights of the final stock prices
x=rbinom(n.sim, n, p)

# simulated stock prices
s.T=s0*u^x*d^(n-x)

# simulated payoffs
v.T=pmax(s.T-K,0) # caveat: it MUST be the pointwise max
#v.T

# average payoff
v.bar=mean(v.T)

# discounting 
(v.0.mc=exp(-r*T)*v.bar)
```

