---
title: "Problem set #13: Black-Scholes Monte Carlo: Solutions"
author: "Milica Cudina"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

Let the continuously compounded, risk-free interest rate be $0.04$. 

```{r}
r=0.04
```

Consider a stock whose current price is \$100 and whose volatility is $0.25$. We will be pricing a one-year, \$95-strike call option. 

```{r}
#about the stock
s0=100
sigma=0.25

#about the call
T=1
K=95
```

## Problem #1: Analytic Black-Scholes
Price the option above using the Black-Scholes pricing formula. 

*Solution:*
```{r}
#this is the call
(d.one=(log(s0/K)+(r+sigma^2/2)*T)/(sigma*sqrt(T)))

(d.two=d.one-sigma*sqrt(T))

(n.d.one=pnorm(d.one))
(n.d.two=pnorm(d.two))

(v.c.0=s0*n.d.one-K*exp(-r*T)*n.d.two)

#just for fun
(v.p.0=K*exp(-r*T)*pnorm(-d.two)-s0*pnorm(-d.one))

#put-call parity
(v.c.0-v.p.0)
(s0-K*exp(-r*T))
(v.c.0-v.p.0-s0+K*exp(-r*T)==0)
```

## Problem #2: Black-Scholes Monte Carlo with $Z$
Provide the *Monte Carlo* estimate of the price using the simulated draws from the standard normal distribution with 10000 simulations.

*Solution:*
```{r}
#number of simulated values
n.sims=10^5

#drawing the simulations from N(0,1)
z.s=rnorm(n.sims)
#hist(z.s)

#simulated stock prices
s.T=s0*exp((r-sigma^2/2)*T+sigma*sqrt(T)*z.s)
#hist(s.T)

#simulated payoffs
v.T=pmax(s.T-K,0)
#hist(v.T)

#average of the simulated payoffs
(v.bar=mean(v.T))

#discounting to get the Monte Carlo price
(v.c.mc=exp(-r*T)*v.bar)
```


## Problem #3: Black-Scholes Monte Carlo with `rlnorm``
Provide the *Monte Carlo* estimate of the price using the simulated draws from the lognormal distribution with 10000 simulations.

*Solution:*
```{r}
#shorthand
mu=r-sigma^2/2

#simulated stock prices
s.T=s0*rlnorm(n.sims, meanlog=mu*T, sdlog=sigma*sqrt(T))
hist(s.T, prob=TRUE, ylim=c(0, dlnorm(100, meanlog=log(s0)+mu*T, sdlog=sigma*sqrt(T))))

#just for laughs - superimpose the lognormal density
curve(dlnorm(x,meanlog=log(s0)+mu*T, sdlog=sigma*sqrt(T)), col="blue", add=TRUE)

#back to pricing
#simulated payoffs
v.T=pmax(s.T-K, 0)

#the Monte Carlo price
(v.mc=exp(-r*T)*mean(v.T))
```

