---
title: "Problem set #13: Black-Scholes Monte Carlo: Solutions"
author: "Milica Cudina"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

Let the continuously compounded, risk-free interest rate be $0.04$. 

```{r}
r=0.04
```


Consider a stock whose current price is \$100 and whose volatility is $0.25$. We will be pricing a one-year, \$95-strike call option. 

```{r}
s0=100
sigma=0.25
T=1
K=95
```


## Problem #1: Analytic Black-Scholes
Price the option above using the Black-Scholes pricing formula. 

*Solution:*
```{r}
d.one=(log(s0/K)+(r+sigma^2/2)*T)/sigma/sqrt(T)
#d.one
d.two=d.one-sigma*sqrt(T)
#d.two

n.d.one=pnorm(d.one)
n.d.two=pnorm(d.two)

v.c=s0*n.d.one-K*exp(-r*T)*n.d.two
v.c
```


## Problem #2: Black-Scholes Monte Carlo with $Z$
Provide the *Monte Carlo* estimate of the price using the simulated draws from the standard normal distribution with 10000 simulations.

*Solution:*
```{r}
n.sims=10000
z.s=rnorm(n.sims)

#the simulated stock prices
s.T=s0*exp((r-sigma^2/2)*T + sigma*sqrt(T)*z.s)
#hist(s.T)

#simulated payoffs
v.T=pmax(s.T-K,0)
v.bar=mean(v.T)

#MC price
v.c.mc=exp(-r*T)*v.bar
v.c.mc
```

## Problem #3: Black-Scholes Monte Carlo with `rlnorm``
Provide the *Monte Carlo* estimate of the price using the simulated draws from the lognormal distribution with 10000 simulations.

*Solution:*
```{r}
mu=r-sigma^2/2

#simulated stock prices
s.T=s0*rlnorm(n.sims, meanlog=mu*T, sdlog=sigma*sqrt(T))
#hist(s.T)

#simulated payoffs
v.T=pmax(s.T-K, 0)
v.bar=mean(v.T)

v.c.mc=exp(-r*T)*v.bar
v.c.mc
```

