---
title: "Problem set #11: Binomial Monte Carlo"
author: "Milica Cudina"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

Let the **volatility** of a stock be the standard deviation of its (continuously compounded) realized return on an annual basis. Then, we can define the up and down factors in the so-called *forward binomial tree* for a **non-dividend-paying** stock as
\begin{equation}
\begin{split}
u & = e^{rh + \sigma\sqrt{h}}\\
d & = e^{rh - \sigma\sqrt{h}}\\
\end{split}
\end{equation}

Let the continuously compounded, risk-free interest rate be $0.04$. 

```{r}
r=0.04
```


Consider a stock whose current price is \$100 and whose volatility is $0.25$. We will be pricing a one-year, at-the-money call option in a variety of ways here. 

```{r}
#about the stock
s0=100
sigma=0.25
#about the call
T=1
K=s0
```


## Problem #1: Analytic one period
Price the option above using a one period binomial tree. 

```{r}
#length of period?
n=1
h=T/n

#factors?
u=exp(r*h+sigma*sqrt(h))
#u
d=exp(r*h-sigma*sqrt(h))
#d

#the risk-neutral probability?
#by definition
p=(exp(r*h)-d)/(u-d)
p
#forward tree shortcut
p.1=1/(1+exp(sigma*sqrt(h)))
p.1

#the possible stock prices
s.u=s0*u
s.d=s0*d

#the possible payoffs
v.u=max(s.u-K,0)
v.u
v.d=max(s.d-K,0)
v.d

#initial option price
v.0=exp(-r*T)*(v.u*p+v.d*(1-p))
v.0
```


## Problem #2: Monte Carlo one period
Price the option above using Monte Carlo a one period binomial tree. Use $10000$ simulations. 

```{r}
#number of simulations
nsims=10^6

#possible results of the cointoss
factors=c(u,d)
#risk-neutral probabilities
probs=c(p, 1-p)

#simulated final stock prices
s.T=s0*sample(factors, size=nsims, prob=probs, replace=TRUE)
#head(s.T)

#simulated payoffs
v.T=pmax(s.T-K,0) #careful about pointwise maximum
#v.T

#expected risk-neutral payoff
v.bar=mean(v.T)
#v.bar

#Monte-Carlo price
v.mc=exp(-r*T)*v.bar
v.mc
```


## Problem #3: Analytic two periods
Price the above option using a two-period binomial tree. 


## Problem #4: Monte Carlo two periods
Price the option above using Monte Carlo a two period binomial tree. Use $10000$ simulations. 


## Problem #5: Analytic one hundred periods
Price the above option using a $100-$period binomial tree. 


## Problem #6: Monte Carlo with one hundred periods
Price the option above using Monte Carlo with a hundred period binomial tree. Use $10000$ simulations. 

