---
title: "Problem set #11: Binomial Monte Carlo"
author: "Milica Cudina"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

Let the **volatility** of a stock be the standard deviation of its (continuously compounded) realized return on an annual basis. Then, we can define the up and down factors in the so-called *forward binomial tree* for a **non-dividend-paying** stock as
\begin{equation}
\begin{split}
u & = e^{rh + \sigma\sqrt{h}}\\
d & = e^{rh - \sigma\sqrt{h}}\\
\end{split}
\end{equation}

Let the continuously compounded, risk-free interest rate be $0.04$. 

```{r}
r=0.04
```

Consider a stock whose current price is \$100 and whose volatility is $0.25$. We will be pricing a one-year, at-the-money call option in a variety of ways here. 

```{r}
#about the stock
s0=100
sigma=0.25

#about the option
T=1
K=s0
```


## Problem #1: Analytic one period
Price the option above using a one period binomial tree. 

```{r}
#number of periods
n=1

#length of each period
h=T/n

#define the factors in the forward binomial tree
u=exp(r*h+sigma*sqrt(h))
d=exp(r*h-sigma*sqrt(h))

(u)
(d)

#the risk-neutral probability
(p=(exp(r*h)-d)/(u-d))

#the forward-tree shortcut
(p.dash=1/(1+exp(sigma*sqrt(h))))

#the possible stock prices
(s.u=s0*u)
(s.d=s0*d)

#the possible payoffs
(v.u=max(s.u-K,0))
(v.d=max(s.d-K,0))

#initial option price
(v0=exp(-r*T)*(v.u*p+v.d*(1-p)))
```



## Problem #2: Monte Carlo one period
Price the option above using Monte Carlo a one period binomial tree. Use $10000$ simulations. 

```{r}
#number of simulations
n.sim=10^4

#possible results of the cointoss
factors=c(u,d)
#their respective risk-neutral probabilities
probs=c(p, 1-p)

#simulated stock prices
s.T=s0*sample(factors, size=n.sim, prob=probs, replace=TRUE)
#(s.T)

#simulated payoffs
(v.T=pmax(s.T-K, 0)) #caveat: it MUST be the pointwise max

#averaging the simulated payoffs
(v.bar=mean(v.T))

#discounting
(v.0.mc=exp(-r*T)*v.bar)

```

## Problem #3: Analytic two periods
Price the above option using a two-period binomial tree. 

```{r}
#number of periods
n=2

#length of each period
h=T/n

#define the factors in the forward binomial tree
u=exp(r*h+sigma*sqrt(h))
d=exp(r*h-sigma*sqrt(h))

(u)
(d)

#the risk-neutral probability
(p=(exp(r*h)-d)/(u-d))

#possible stock prices
(s.T=c(s0*u^2, s0*u*d, s0*d^2))

#possible payoffs
(v.T=pmax(s.T-K,0))

#risk-neutral probabilities of reaching the terminal nodes
(probs=dbinom(c(2,1,0), size=2, prob=p))

#expected risk-neutral payoff
(sum(v.T*probs))

#discounting
(v.0=exp(-r*T)*sum(v.T*probs))
```

## Problem #4: Monte Carlo two periods
Price the option above using Monte Carlo a two period binomial tree. Use $10000$ simulations. 

```{r}
#steppingstone: 
#simulating the heights of the terminal stock prices
(x=rbinom(n.sim, size=2, prob=p))
#the simulated stock prices
(s.T=s0*u^x*d^(n-x))

#simulated payoffs
(v.T=pmax(s.T-K, 0)) #caveat: it MUST be the pointwise max

#averaging the simulated payoffs
(v.bar=mean(v.T))

#discounting
(v.0.mc=exp(-r*T)*v.bar)
```

## Problem #5: Analytic one hundred periods
Price the above option using a $100-$period binomial tree. 

```{r}
#number of periods
n=100

#length of each period
h=T/n

#define the factors in the forward binomial tree
u=exp(r*h+sigma*sqrt(h))
d=exp(r*h-sigma*sqrt(h))

(u)
(d)

#the risk-neutral probability
(p=(exp(r*h)-d)/(u-d))

#possible stock prices
k=n:0
(s.T=s0*u^k*d^(n-k))

#possible payoffs
(v.T=pmax(s.T-K,0))

#their respective probabilities
(probs=dbinom(k, size=n, prob=p))

#option price
(v.0=exp(-r*T)*sum(v.T*probs))
```


## Problem #6: Monte Carlo with one hundred periods
Price the option above using Monte Carlo with a hundred period binomial tree. Use $10000$ simulations. 

```{r}
#steppingstone: 
#simulating the heights of the terminal stock prices
(x=rbinom(n.sim, size=100, prob=p))
#the simulated stock prices
(s.T=s0*u^x*d^(n-x))

#simulated payoffs
(v.T=pmax(s.T-K, 0)) #caveat: it MUST be the pointwise max

#averaging the simulated payoffs
(v.bar=mean(v.T))

#discounting
(v.0.mc=exp(-r*T)*v.bar)
```

