---
title: "Problem set #8: Binomial Monte Carlo: Solutions"
author: "Milica Cudina"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

Let the **volatility** of a stock be the standard deviation of its (continuously compounded) realized return on an annual basis. Then, we can define the up and down factors in the so-called *forward binomial tree* for a **non-dividend-paying** stock as
\begin{equation}
\begin{split}
u & = e^{rh + \sigma\sqrt{h}}\\
d & = e^{rh - \sigma\sqrt{h}}\\
\end{split}
\end{equation}

Let the continuously compounded, risk-free interest rate be $0.04$. 
```{r}
r=0.04
```

Consider a non-dividend-paying stock whose current price is \$100 and whose volatility is $0.25$. We will be pricing a one-year, at-the-money call option in a variety of ways here. 
```{r}
s0=100
sigma=0.25
T=1
K=s0
```


## Problem #1: Analytic one period
Price the option above using a one period binomial tree. 

*Solution:*
```{r}
periods=1
h=T/periods
u=exp(r*h+sigma*sqrt(h))
d=exp(r*h-sigma*sqrt(h))
u
d

#the payoff function
v.c<-function(x){
  max(x-K,0)
}

p.star=(exp(r*h)-d)/(u-d)
p.star

#the possible stock prices
s.u=s0*u
s.d=s0*d

#the possible payoffs
v.u=v.c(s.u)
v.d=v.c(s.d)
  
v.0=exp(-r*T)*(p.star*v.u+(1-p.star)*v.d)
v.0
```

## Problem #2: Monte Carlo one period
Price the option above using Monte Carlo a one period binomial tree. Use $10000$ simulations. 

*Solution:*
```{r}
n.sims=10000
probs=c(p.star, 1-p.star)
factors=c(u,d)
final.prices=s0*sample(factors, size=n.sims, prob=probs, replace=TRUE)
#final.prices
payoffs<-pmax(final.prices-K,0)
#payoffs
v.bar=mean(payoffs)
#v.bar
v.0.mc<-exp(-r*T)*v.bar
v.0.mc
```

## Problem #3: Analytic two periods
Price the above option using a two-period binomial tree. 

*Solution:*
```{r}
periods=2
h=T/periods
#h
u=exp(r*h+sigma*sqrt(h))
d=exp(r*h-sigma*sqrt(h))
#u
#d

p.star=(exp(r*h)-d)/(u-d)
#p.star

s.T=s0*c(u^2, u*d, d^2)
#s.T

v.T=pmax(s.T-K,0)
#v.T

probs=c(p.star^2, 2*p.star*(1-p.star), (1-p.star)^2)

v.c=exp(-r*T)*sum(v.T*probs)
v.c
```


## Problem #4: Monte Carlo two periods
Price the option above using Monte Carlo using a two period binomial tree. Use $10000$ simulations. 

*Solution:*
```{r}
#get the simulated values of the final stock price
#it's sufficient to simulate the number of upsteps
x=rbinom(n.sims, size=periods, prob=p.star)
x
s.T=s0*u^x*d^(periods-x)
s.T

#get the payoffs
v.T=pmax(s.T-K,0)

v.bar=mean(v.T)
v.c.mc=exp(-r*T)*v.bar
v.c.mc
```

## Problem #5: Analytic one hundred periods
Price the above option using a $100-$period binomial tree. 

*Solution:*
```{r}
periods=100
h=T/periods
#h
u=exp(r*h+sigma*sqrt(h))
d=exp(r*h-sigma*sqrt(h))
#u
#d

p.star=(exp(r*h)-d)/(u-d)
#p.star

k=periods:0
s.T=s0*u^k*d^(periods-k)
s.T

v.T=pmax(s.T-K, 0)

rn.probs=dbinom(k, size=periods, prob=p.star)

v.c=exp(-r*T)*sum(v.T*rn.probs)
v.c
```

## Problem #6: Monte Carlo with one hundred periods
Price the option above using Monte Carlo with a hundred period binomial tree. Use $10000$ simulations. 

*Solution:*
```{r}
#simulating the number of upsteps
x=rbinom(n.sims, size=periods, prob=p.star)
#x

#calculate the corresponding simulated stock prices
s.T=s0*u^x*d^(periods-x)
#hist(s.T, breaks=25)

#calculating the simulated payoffs
v.T=pmax(s.T-K,0)
#v.T

#average the simulated payoffs
v.bar=mean(v.T)
#v.bar

#discounting
v.0.mc=exp(-r*T)*v.bar
v.0.mc
```


