---
title: "Problem set #10: Binomial Monte Carlo"
author: "Milica Cudina"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

Let the **volatility** of a stock be the standard deviation of its (continuously compounded) realized return on an annual basis. Then, we can define the up and down factors in the so-called *forward binomial tree* for a **non-dividend-paying** stock as
\begin{equation}
\begin{split}
u & = e^{rh + \sigma\sqrt{h}}\\
d & = e^{rh - \sigma\sqrt{h}}\\
\end{split}
\end{equation}

Let the continuously compounded, risk-free interest rate be $0.04$. 
```{r}
r=0.04
```


Consider a stock whose current price is \$100 and whose volatility is $0.25$. We will be pricing a one-year, at-the-money call option in a variety of ways here. 

```{r}
#about the stock
s0=100
sigma=0.25
#about the call 
T=1
K=s0
```


## Problem #1: Analytic one period
Price the option above using a one period binomial tree. 

```{r}
h=1
u.1=exp(r*h+sigma*sqrt(h))
#u.1
d.1=exp(r*h-sigma*sqrt(h))
#d.1

s.u=s0*u.1
s.d=s0*d.1

v.c<-function(x){
  max(x-K,0)
}

p.star=1/(1+exp(sigma*sqrt(h)))
  
v.0=exp(-r*T)*(v.c(s.u)*p.star+v.c(s.d)*(1-p.star))
v.0
```


## Problem #2: Monte Carlo one period
Price the option above using Monte Carlo a one period binomial tree. Use $10000$ simulations. 

```{r}
nsims=10000

probs=c(p.star, 1-p.star)
factors=c(u.1, d.1)

s.T=s0*sample(factors, size=nsims, prob=probs, replace=TRUE)
#s.T

v.T=pmax(s.T-K, 0)
#v.T

v.bar=mean(v.T)
#v.bar

v.0.mc=exp(-r*T)*v.bar
v.0.mc
```

## Problem #3: Analytic two periods
Price the above option using a two-period binomial tree. 

```{r}
n=2
h=T/n

u.2=exp(r*h+sigma*sqrt(h))
#u.2
d.2=exp(r*h-sigma*sqrt(h))
#d.2

#possible stock prices
s.T=s0*c(u.2^2, u.2*d.2, d.2^2)
#s.T

#the risk-neutral probability
p.star=(exp(r*h)-d.2)/(u.2-d.2)
#p.star

#just for laughs, in the the FORWARD binomial tree, we also have
p.tilde=1/(1+exp(sigma*sqrt(h)))
#p.tilde

#the possible payoffs are
v.T=pmax(s.T-K, 0)
#v.T

#the probabilities in the final nodes
probs=c(p.star^2, 2*p.star*(1-p.star), (1-p.star)^2)
#probs

#the risk-neutral price
v.0=exp(-r*T)*sum(v.T*probs)
v.0
```


## Problem #4: Monte Carlo two periods
Price the option above using Monte Carlo a two period binomial tree. Use $10000$ simulations. 

```{r}
nsims=10000

#simulated number of upsteps
x=rbinom(nsims, size=2, prob=p.star)
#x

#simulated final stock prices
s.T=s0*u.2^x*d.2^(n-x)
#s.T

#applying the payoff function
v.T=pmax(s.T-K,0)
#v.T

#Monte Carlo estimate of the risk-neutral expectation
v.bar=mean(v.T)

#MC price
v.0.mc=exp(-r*T)*v.bar
v.0.mc
```


## Problem #5: Analytic one hundred periods
Price the above option using a $100-$period binomial tree. 

```{r}
n=100
h=T/n

#up and down factors
u.n=exp(r*h+sigma*sqrt(h))
#u.n
d.n=exp(r*h-sigma*sqrt(h))
#d.n

#risk-neutral probability
p.star=(exp(r*h)-d.n)/(u.n-d.n)
#p.star

#the possible final stock price
k=n:0
s.T=s0*u.n^k*d.n^(n-k)
#s.T

#the possible payoffs
v.T=pmax(s.T-K,0)
#v.T

#the risk-neutral probabilities of reaching the possible payoffs
rn.probs=dbinom(k, size=n, prob=p.star)

#the option price
v.0=exp(-r*T)*sum(v.T*rn.probs)
v.0
```

## Problem #6: Monte Carlo with one hundred periods
Price the option above using Monte Carlo with a hundred period binomial tree. Use $10000$ simulations. 

```{r}
nsims=10000

#simulated numbers of upsteps
x=rbinom(nsims, size=n, prob=p.star)

#simulated final stock prices
s.T=s0*u.n^x*d.n^(n-x)

#the resulting simulated payoffs
v.T=pmax(s.T-K, 0)

v.bar=mean(v.T)

v.0.mc=exp(-r*T)*v.bar
v.0.mc
```

