---
title: "Problem set #7: Monte Carlo"
author: "Milica Cudina"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

## Problem #1: Simple gambling
Consider a game in which a spinner is spun. The spinner is equally likely to land on *yellow*, *red*, or *blue*. If the spinner lands on *yellow*, the player loses \$1. If the spinner lands on *red*, the player wins \$1. If the spinner lands on *blue*, no money changes hands. 

Use *Monte Carlo* to estimate the expected winnings based on 10,000 runs. Does the result agree with your theoretical expected value?

Now, complete 1,000 batches of 10,000 runs. What does the histogram of the averages of the 1,000 batches look like? What is their mean? What is their variance? What is their standard deviation?

*Solution:*
Let's encode *yellow* as $-1$, *red* as $1$, and *blue* as $0$. 
Let's implement a single run:
```{r}
spinner=c(-1, 0, 1)
one.run=sample(spinner, 1)
one.run
```
We want many runs, so we'll increase the number of simulated values in the `sample` command:
```{r}
nsims=10000
many.runs=sample(spinner, nsims, replace=TRUE)
tab=table(many.runs)
tab
barplot(tab, col="peru")
```
Let's calculate the average of the simulated values
```{r}
mean(many.runs)
```
Create a function which calculates the average winnings for each batch of spins.
```{r}
spin.mc<-function(){
  return(mean(sample(spinner, nsims, replace=TRUE)))
}
spin.mc()
```
We repeat the same function $1000$ times. 
```{r}
num.batches=1000
all.batches=replicate(num.batches, spin.mc())
hist(all.batches, breaks=18)
mean(all.batches)
var(all.batches)
sd(all.batches)
```
What's the theoretical variance of a single spin's winnings?
```{r}
var=1^2*(1/3)+0^2*(1/3)+(-1)^2*(1/3)
var.sample.mean=var/(10000)
var.sample.mean
```



## Problem #2: Gambling
Consider the following game: 

- if a fair die lands on an odd number, the player loses one dollar; 
- if a fair die lands on an even number which is not a perfect square, no money changes hands; 
- if a fair die lands on an even number which is also a perfect square, the player wins nine dollars. 

What is the theoretical expectation of the amount won in this game?

Use *Monte Carlo* to estimate the expected winnings based on 10,000 runs. Does the result agree with your theoretical expected value?

*Solution:*
The probability of losing one dollar is $1/2$. The probability of winning nine dollars $1/6$. The probability of nothing is $1/3$. 
```{r}
p.loss=1/2
p.win=1/6
p.no=1/3
prob=c(1/2, 1/6, 1/3)
win=c(-1, 9, 0)
exp.game=sum(prob*win)
exp.game
```
Using Monte Carlo, we do this:
```{r}
nsims=10000
many.gambles=sample(win, nsims, replace=TRUE, prob=prob)
avg.gamble=mean(many.gambles)
avg.gamble
```


## Problem #3: The binomial distribution
Consider a random variable $X$ which has the binomial distribution with the number of trials equal to $50$ and the probability of success in every trial equal to $2/5$. 

What is the expected value of this random variable?

Use *Monte Carlo* to estimate the expected value based on 10,000 runs. Does the result agree with your theoretical expected value?

*Solution:*
Specifying the parameters:
```{r}
n.trials=50
prob.s=2/5
```
Theoretical (exact) expected value
```{r}
exp.X=n.trials*prob.s
exp.X
```

Using Monte Carlo, we get:
```{r}
runs<-rbinom(10000, size=50, prob=2/5)
mean(runs)
```

