---
title: "Bagging and Random Forests"
author: "Trevor Hastie and Robert Tibshirani"
output:
  pdf_document: default
---

**Here, I am adapting part of the lab associated with Chapter 8 of the textbook.** 

The `tree` library is used to construct classification and regression trees.

```{r chunk1}
#install.packages("tree")
library(tree)
library(ISLR2)
```

## Bagging and Random Forests
Here we apply bagging and random forests to the `Boston` data, using the `randomForest` package in `R`. The exact results obtained in this section may depend on the version of `R` and the version of the `randomForest` package installed on your computer.

We start by splitting the data into training and testing. 

```{r}
set.seed(1)
train <- sample(1:nrow(Boston), nrow(Boston) / 2)
```

Recall that bagging is simply a special case of a random forest with
$m=p$. Therefore, the
`randomForest()` function can be used to perform both random forests and bagging.
We perform bagging as follows:

```{r chunk19}
#install.packages("randomForest")
library(randomForest)

bag.boston <- randomForest(medv ~ ., data = Boston,
    subset = train, mtry = 12, importance = TRUE)
bag.boston
importance(bag.boston)
```

The argument `mtry = 12` indicates that all $12$ predictors should be considered for each split of the tree---in other words, that bagging should be done.
How well does this bagged model perform on the test set?

```{r chunk20}
yhat.bag <- predict(bag.boston, newdata = Boston[-train, ])
boston.test <- Boston[-train, "medv"]

plot(yhat.bag, boston.test, pch=19, col="lightblue")
abline(0, 1, col="blue")
mean((yhat.bag - boston.test)^2)
```

The test set MSE associated with the bagged regression tree is $23.42$, about two-thirds of that was obtained using an optimally-pruned single tree (we got $35.28688$ last week).
We could change the number of trees grown by `randomForest()` using the `ntree` argument:

```{r chunk21}
bag.boston <- randomForest(medv ~ ., data = Boston,
    subset = train, mtry = 12, ntree = 25)
yhat.bag <- predict(bag.boston, newdata = Boston[-train, ])
mean((yhat.bag - boston.test)^2)
```


Growing a random forest proceeds in exactly the same way, except that we use a smaller value of the `mtry` argument. By default, `randomForest()` uses $p/3$ variables when building a random forest of regression trees, and $\sqrt{p}$ variables when building a random forest of classification trees. Here we use `mtry = 6`.

```{r chunk22}
set.seed(1)
rf.boston <- randomForest(medv ~ ., data = Boston,
    subset = train, mtry = 6, importance = TRUE)
yhat.rf <- predict(rf.boston, newdata = Boston[-train, ])
mean((yhat.rf - boston.test)^2)
```

The test set MSE is $20.07$; this indicates that random forests yielded an improvement over bagging in this case.

Using the `importance()` function, we can view the importance of each variable.

```{r chunk23}
importance(rf.boston)
```

Two measures of variable importance are reported. The first is based upon the mean decrease of accuracy in predictions on the out of bag samples when a given variable is permuted. The second is a measure of the total decrease in node impurity that results from splits over that variable, averaged over all trees. In the case of regression trees, the node impurity is measured by the training RSS, and for classification trees by the deviance. Plots of these importance measures can be produced using the `varImpPlot()` function.

```{r chunk24}
varImpPlot(rf.boston)
```

The results indicate that across all of the trees considered in the random forest, the wealth of the community (`lstat`) and the house size (`rm`) are by far the two most important variables.
