---
title: "LDA: Banknotes"
author: "Gustavo Cepparo and Milica Cudina"
output:
  pdf_document: default
---


We reconsider the **banknote** data set. It exists in the `MixGHD` library. 

```{r}
#install.packages("MixGHD")
library(MixGHD)
```

The data set `banknote` is used to predict the `Status` of Swiss Franc bills based on measurements of the bills. 

```{r}
data(banknote)
names(banknote)
attach(banknote)
data<-banknote
```

We can start with visualization (as usual). 

```{r}
pairs(data, 
      col="lightblue", pch=19)
```

Our aim is to perform the linear discriminant analysis here. First, we will partition our data into training and testing. 

```{r}
set.seed(1)
ind.train <- sample(nrow(data), floor(nrow(data)*0.6))
training<-data[ind.train,]
testing<-data[-ind.train,]
```

My goal with this example is to ultimately demonstrate how to plot a partition in the plane. So, I will arbitrarily choose two explanatory variables: `Top` and `Diagonal`.

```{r}
library(MASS)
linear <- lda(Status~Top+Diagonal, data=training)
summary(linear)
```
# Histogram
```{r}
p <- predict(linear, training)
ldahist(data = p$x[,1], g = training$Status)
```


