---
title: "Linear Discriminant Analysis (LDA)"
author: "Trevor Hastie and Robert Tibshirani"
output:
  pdf_document: default
---

**Here, I am adapting part of the lab associated with Chapter 4 of the textbook.** 

We re-examine the `Smarket` data, which is part of the `ISLR2` library. This data set consists of percentage returns for the S\&P 500 stock index over $1,250$ days, from the beginning of 2001 until the end of 2005. For each date, we have recorded the percentage returns for each of the five previous trading days, `lagone` through `lagfive`. We have also recorded `volume` (the number of shares traded on the previous day, in billions), `Today` (the percentage return on the date in question)  and `direction` (whether the market was `Up` or `Down` on this date). Our goal is to predict `direction` (a qualitative response) using the other features.

```{r chunk1}
library(ISLR2)
names(Smarket)
dim(Smarket)
summary(Smarket)
attach(Smarket)
```

We will again create a vector corresponding to the observations from 2001 through 2004. We will then use this vector to create a held out data set of observations from 2005.

```{r chunk9}
train <- (Year < 2005)
Smarket.2005 <- Smarket[!train, ]
dim(Smarket.2005)
Direction.2005 <- Direction[!train]
```

## Linear Discriminant Analysis

Now we will perform LDA on the `Smarket` data. In `R`, we fit an LDA model using the  `lda()` function, which is part of the `MASS` library. Notice that the syntax for the `lda()` function is identical to that of `lm()`, and to that of `glm()` except for the absence of the `family` option. We fit the model using only the observations before 2005.

```{r chunk14}
library(MASS)
lda.fit <- lda(Direction ~ Lag1 + Lag2, data = Smarket,
    subset = train)
lda.fit
plot(lda.fit)
```

The LDA output indicates that $\hat\pi_1=0.492$ and $\hat\pi_2=0.508$; in other words, $49.2$ \% of the training observations correspond to days during which the market went down.
It also provides the group means; these are the average of each predictor within each class, and are used by LDA as estimates of $\mu_k$. These suggest that there is a tendency for the previous 2 days' returns to be negative on days when the market increases, and a tendency for the previous days' returns to be positive on days when the market declines. The *coefficients of linear discriminants* output provides the linear combination of `lagone` and `lagtwo` that are used to form the LDA decision rule. In other words, these are the multipliers of the elements of $X=x$ in 
$$
\delta_k(x) = x^T\mathbf{\Sigma}^{-1}\mu_k - \frac{1}{2}\mu_k^T\mathbf{\Sigma}^{-1}\mu_k + \log(\pi_k)
$$
once the expression is simplified. 
If $-0.642 \times \texttt{`lagone`} - 0.514 \times \texttt{`lagtwo`}$ is large, then the LDA classifier will predict a market increase, and if it is small, then the LDA classifier will predict a market decline.

The `plot()` function produces plots of the *linear discriminants*, obtained by computing $-0.642 \times \texttt{`lagone`} - 0.514 \times \texttt{`lagtwo`}$ for each of the training observations. The `Up` and `Down` observations are displayed separately.


The `predict()` function returns a list with three elements. The first element,  `class`,  contains LDA's predictions about the movement of the market. The second element, `posterior`, is a matrix whose $k$th column contains the posterior probability that the corresponding observation belongs to the $k$th class, computed from 
$$
\mathbb{P}[Y =k \,|\, X=x] 
= \frac{\pi_kf_k(x)}{\sum_{j=1}^K \pi_j f_j(x)}
$$

Finally, `x` contains the linear discriminants, described earlier.

```{r chunk15}
lda.pred <- predict(lda.fit, Smarket.2005)
names(lda.pred)
```

As we observed in Section 4.5, the LDA and logistic regression predictions are almost identical.

```{r chunk16}
lda.class <- lda.pred$class
table(lda.class, Direction.2005)
mean(lda.class == Direction.2005)
```

Applying a $50$ \% threshold to the posterior probabilities allows us to recreate the predictions contained in `lda.pred$class`.

```{r chunk17}
sum(lda.pred$posterior[, 1] >= .5)
sum(lda.pred$posterior[, 1] < .5)
```

Notice that the posterior probability output by the model corresponds to the probability that the market will *decrease*:

```{r chunk18}
lda.pred$posterior[1:20, 1]
lda.class[1:20]
```

If we wanted to use a posterior probability threshold other than $50$ \% in order to make predictions, then we could easily do so. For instance, suppose that we wish to predict a market decrease only if we are very certain that the market will indeed decrease on that day---say, if the posterior probability
is at least $90$ \%.

```{r chunk19}
sum(lda.pred$posterior[, 1] > .9)
```

No days in 2005 meet that threshold! In fact, the greatest posterior probability of decrease in all of 2005 was $52.02$ \%.

## Quadratic Discriminant Analysis

We will now fit a QDA model to the `Smarket` data. QDA is  implemented in `R` using the `qda()` function, which is also part of the `MASS` library. The syntax is identical to that of `lda()`.

```{r chunk20}
qda.fit <- qda(Direction ~ Lag1 + Lag2, data = Smarket,
    subset = train)
qda.fit
```

The output  contains the group means. But it does not contain the coefficients of the linear discriminants, because the QDA classifier involves  a quadratic, rather than a linear, function of the predictors. The `predict()` function works in exactly the same fashion as for LDA.

```{r chunk21}
qda.class <- predict(qda.fit, Smarket.2005)$class
table(qda.class, Direction.2005)
mean(qda.class == Direction.2005)
```

Interestingly, the QDA predictions are accurate almost $60$ \% of the time, even though the 2005 data was not used  to fit the model. This level of accuracy is quite impressive for stock market data, which is known to be quite hard to model accurately.
This  suggests that the quadratic form assumed by QDA may capture the true relationship more accurately than the linear forms assumed by LDA and logistic regression.
 However, we recommend evaluating this method's performance on a larger test set  before betting  that this approach will consistently beat the market!

## Naive Bayes

Next, we fit a naive Bayes model to the `Smarket` data. Naive Bayes is implemented in `R` using the `naiveBayes()` function, which is part of the `e1071` library. The syntax is identical to that of `lda()` and `qda()`.
By default, this implementation of the naive Bayes classifier models each quantitative feature using a Gaussian distribution. However, a kernel density method can also be used to estimate the distributions.

```{r chunk22}
library(e1071)
nb.fit <- naiveBayes(Direction ~ Lag1 + Lag2, data = Smarket,
    subset = train)
nb.fit
```

The output  contains the estimated mean and standard deviation for each variable in each class. For example, the mean for `lagone` is $0.0428$ for   `Direction=Down`, and the standard deviation is $1.23$. We can easily verify this:

```{r chunk23}
mean(Lag1[train][Direction[train] == "Down"])
sd(Lag1[train][Direction[train] == "Down"])
```

The `predict()` function is straightforward.

```{r chunk24}
nb.class <- predict(nb.fit, Smarket.2005)
table(nb.class, Direction.2005)
mean(nb.class == Direction.2005)
```

Naive Bayes performs very well on this data, with accurate predictions over $59\%$ of the time. This is slightly worse than QDA, but much better than LDA.

The `predict()` function can also generate estimates of the probability that each observation belongs to a particular class. 

```{r chunk25}
nb.preds <- predict(nb.fit, Smarket.2005, type = "raw")
nb.preds[1:5, ]
```
