---
title: "Splines"
author: "Trevor Hastie and Robert Tibshirani"
output:
  pdf_document: default
---

Here, I am adapting part of the lab associated with Chapter 7 of the textbook. 

In this lab, we re-analyze the `Wage` data considered in the
examples throughout this chapter, in order to illustrate the fact that many of the complex non-linear fitting procedures discussed can be easily implemented in `R`.  We begin by loading the `ISLR2` library, which contains the data.

```{r chunk1}
library(ISLR2)
attach(Wage)
dim(Wage)
plot(age, wage,
     pch=20, col="lightblue")
```

**I am inserting some more exploratory data analysis here based on student feedback.**

```{r chunk1a}
library(ggplot2)

df=data.frame(age,wage)

ggplot(df, aes(x = age, y = wage)) +
  geom_point(color="lightblue") +
  stat_ellipse(level = 0.9) +
  stat_ellipse(level = 0.95, color = 2) +
  stat_ellipse(level = 0.99, color = 3)
```


## Cubic splines

In order to fit regression splines in `R`, we use the `splines` library. In Section 7.4, we saw that regression
splines can be fit by constructing an appropriate matrix of basis functions. The `bs()` function generates the entire matrix of basis functions for splines with the specified set of knots. By default, cubic splines are produced.
Fitting `wage` to `age` using a regression spline is simple:

```{r chunk19}
library(splines)
fit <- lm(wage ~ bs(age, knots = c(25, 40, 60)), data = Wage)

agelims <- range(age)
age.grid <- seq(from = agelims[1], to = agelims[2])
preds <- predict(fit, newdata = list(age = age.grid),
    se = TRUE)
se.bands <- cbind(preds$fit + 2 * preds$se.fit,
    preds$fit - 2 * preds$se.fit)

pred <- predict(fit, newdata = list(age = age.grid), se = T)
plot(age, wage, col = "gray", pch=20)
lines(age.grid, pred$fit, lwd = 2)
lines(age.grid, pred$fit + 2 * pred$se, lty = "dashed", col="blue")
lines(age.grid, pred$fit - 2 * pred$se, lty = "dashed", col="blue")
```

Here we have prespecified knots at ages $25$, $40$, and $60$. This produces a spline with six basis functions. (Recall that a cubic spline with three knots has seven degrees of freedom; these degrees of freedom are used up by an intercept, plus six basis functions.) We could also use the `df` option to produce a spline with knots at uniform quantiles
of the data.

```{r chunk20}
dim(bs(age, knots = c(25, 40, 60)))
dim(bs(age, df = 6))
attr(bs(age, df = 6), "knots")
```

In this case `R` chooses knots at ages $33.8, 42.0$, and $51.0$, which correspond to the 25th, 50th, and 75th percentiles of `age`. The function `bs()` also has a `degree` argument, so we can fit splines of any degree, rather than the default degree of 3 (which yields a cubic spline). 

In order to instead fit a natural spline, we use the `ns()` function.
Here we fit a natural spline with four degrees of freedom.

```{r chunk21}
fit2 <- lm(wage ~ ns(age, df = 4), data = Wage)
pred2 <- predict(fit2, newdata = list(age = age.grid),
     se = T)
plot(age, wage, col = "gray", pch=20)
lines(age.grid, pred2$fit, col = "red", lwd = 2)
lines(age.grid, pred$fit, lwd = 2, col="blue")
```

As with the `bs()` function, we could instead specify the knots directly using the `knots` option.

In order to fit a smoothing spline, we use
the `smooth.spline()` function.
Figure 7.8  was produced with the following
code:

```{r chunk22}
plot(age, wage, xlim = agelims, cex = .5, col = "darkgrey",
     pch=20)
title("Smoothing Spline")
fit <- smooth.spline(age, wage, df = 16)
fit2 <- smooth.spline(age, wage, cv = TRUE)
fit2
fit2$df
lines(fit, col = "red", lwd = 2)
lines(fit2, col = "blue", lwd = 2)
legend("topright", legend = c("16 DF", "6.8 DF"),
    col = c("red", "blue"), lty = 1, lwd = 2, cex = .8)
```

Notice that in the first call to `smooth.spline()`, we specified `df = 16`. The function then  determines which value of $\lambda$  leads to $16$ degrees of freedom.
 In the second call to `smooth.spline()`, we select the smoothness level by cross-validation;
 this results in a value of $\lambda$ that yields 6.8 degrees of freedom.

** How do we create confidence intervals here? In this case, the preferred technique is to generate Bayesian "confidence intervals" using bootstrap**



** Now, here is the last spline in `ggplot2` with "confidence intervals". **
```{r chunk22a}
library(ggformula)

ggplot(df, aes(x = age, y = wage)) +
  geom_point(color="lightblue") +
  stat_smooth(color="blue") 
```
