---
title: "Project #3"
author: "Milica Cudina"
date: "`r Sys.Date()`"
output: pdf_document
urlcolor: blue
---
<!-- The author of this template is Dr. Gordan Zitkovic.-->
<!-- The code chunk below contains some settings that will  -->
<!-- make your R code look better in the output pdf file.  -->
<!-- If you are curious about what each option below does, -->
<!-- go to https://yihui.org/knitr/options/ -->
<!-- If not, feel free to disregard everything ...  -->
```{r echo=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.align="center",
  fig.pos="t",
  strip.white = TRUE
)
```
<!-- ... down to here. -->

---

# Problems

## Problem #1 **(60 points)**
The `iris` data set is built-in in `R`. Start by studying the documentation of the data set, i.e., by entering `?iris` in the console. To familiarize yourselves with the architecture of an iris flower, go to: 

[US Forest Service](https://www.fs.usda.gov/wildflowers/beauty/iris/flower.shtml)

Your next step is exploratory data analysis. 

**(10 points)**
Which plot would you use to display pairwise associations between different measurements? How do you make sure that the different species are color-coded? Display the plot and write a few sentences about your conclusions. 

### Principal Component Analysis (PCA)
**($20$ points)**
Perform the PCA on the explanatory components of the above data, provide the report, and the relevant plots. 

### $K-$Means Clustering After PCA
**($5$ points)**
Using only the first two principal components, perform $K-$means clustering using $K=3$. Provide the plot and report your conclusions.  

### Principal Components Regression (PCR)
Your next task is to predict `Sepal.Length` using the other applicable variables in the `iris` dataset. 

**(15 points)**
Run the PCR, provide an explanation for the output, and display the relevant plots (both validation and prediction). 

**(10 points)**
Split your dataset into training ($4/5$ of the data) and testing ($1/5$ of the data). Provide the mean squared error and an appropriate plot. 

## Problem #2 **(2+5+3+5+5+5+5+5+5+10=50 points)**
Solve **Problem 4.8.13** (pp. 192-193) from the textbook. 

*Hint:*  Here is a list of libraries you will need: 

```{r}
library(MASS)
library(ISLR2)
library(e1071)
library(class)
```


